---
title: "R Notebook"
output: html_notebook
---

Created by Jack Petersen.

## Introduction:

This notebook will take an existing code framework used for processing data and adapt it to 2 slightly different datasets. The goal is to show how to make a new data processing function for plate reader assays by only making slight changes to our existing code. These case studies will focus on processing the data into easily usable files (a single UV spectrum). To make things simpler, we will only be save the files locally and will not be uploading them to the Shared Google Drive folder.

Original Code: survey1_process_data.R

Dataset 1: 1B-pH Plate.xlsx

Dataset 2: 1G+4B - pH Plate.xlsx

NOTE: Original code is included at the bottom

## PACKAGES USED:

run this code to install all of the packages used in this notebook if you have not done so already:

```{r}
install.packages("readxl")
install.packages("writexl")
```

## CASE STUDY 1: 1B-pH Plate.xlsx

```{r}
library(readxl)
library(writexl)

process_data1 <- function(data_path) {
  
  #I have deleted the avg_data() function because these experiments are NOT being run in triplicate, so we are no longer averaging columns in the pH or salt assays (compare to original code)
    

  #---------------------------DEFINE VARIABLES and PREPROCESSING THE DATA------------------
  #define a file name to allow for easy naming of the files later on 
  data_file_name <- "1B_"
  
  #load data file as a .xlsx file using the readxl library. the range gets rid of all of the metadata in the header of the data file 
  data <- read_excel(data_path, range = "1B - pH Plate!A3:CT129") 

  #delete the temperature column 
  data <- data[, -which(names(data) == "Temperature(¡C)")]

  #delete columns with all NA values for all rows (delete empty wells)
  data <- data[, colSums(!is.na(data)) > 0]

  #define how many columns there are using the ncol() function after deleteing all the empty colmns. Will be used for the for loop to create files. 
  num_columns <- ncol(data)

  #extract wavelength column 
  wavelength <- data$Wavelength
  
  #create a mapping of first letters to labels
  label_mapping <- setNames(c('0mM', '200mM', '500mM'), c('A', 'B', 'C'))
  
  #create mapping for pH values based off of second number in column title 
  ph_mapping <- setNames(c('4', '5', '6','7', '8', '9'), c('1', '3', '5', '7', '9', '11'))


  #------------------------------------FOR LOOP-------------------------------------------------

  #There are too many data points to pull cell by cell. Because we know that each column is formatted the same (its a UV spectrum from 450-700nm for that particular cell), we can use that to find the columns we want (non empty cells) then combine them in a new file with our wavelength numbers to create individual spectra for each sample tested 

  #iterate over the columns starting from the second one
  for (i in 2:num_columns) {
    
    #get the column name to determine the label.  
    col_name <- names(data)[i]
    
    #create a new dataframe with the current column as absorbance (wavelength define above)
    current_data <- data.frame(Wavelength = wavelength, data[[i]])
    
    #create variable first_letter to store the first letter of the column name 
    first_letter <- substr(col_name, 1, 1)
    
    #create variable first_letter to store the second letter of the column name 
    second_letter <- substr(col_name, 2, 3)
  
    #get the label based on the first letter
    label <- ifelse(first_letter %in% names(label_mapping), label_mapping[[first_letter]], NA)
  
    #get pH from second letter 
    pH <- ifelse(second_letter %in% names(ph_mapping), ph_mapping[[second_letter]], NA)
    
    
    #construct the file name using the dynamic label
    file_name <- paste0(data_file_name, label, "_NaCl_pH_", pH ,"_Phosphate Buffer", ".xlsx")

    
    #create dataframe and save as .xlsx 
    current_data <- data.frame(Wavelength = wavelength, Absorbance = data[[i]])
    
    #write the file using the writexl library 
    write_xlsx(current_data, path = file_name)
    
    
    #console output 
    message("File: ", file_name, " has been created.")
  
    
  } #END FOR LOOP 

} #END FUNCTION 
```

Run the script:

```{r}
#THIS SCRIPT WILL SAVE FILES TO YOUR WORKING DIRECTORY 
#MAKE SURE TO RUN THE ABOVE CELL BEFORE TRYING TO RUN THIS CELL 


#MAKE SURE TO CHANE THIS FILE PATH TO YOUR LOCAL DATA FILE PATH 
process_data1('/Users/jackpetersen/Downloads/1B - pH Plate.xlsx') 
```

## CASE STUDY 2: 1G+4B - pH Plate.xlsx

For this file, there are 2 separate pH surveys, so we basically have to do our for loop from above on two different parts of the plate.

```{r}
library(readxl)
library(writexl)

process_data2 <- function(data_path) {

  #---------------------DEFINE VARIABLES and PREPROCESSING THE DATA---------------------------
  #load data file as a .xlsx file using the readxl library
  data <- read_excel(data_path, range = "Sheet1!A3:CT129") 
  
  #delete the temperature column 
  data <- data[, -which(names(data) == "Temperature(¡C)")]

  #delete columns with all NA values for all rows (delete empty wells)
  data <- data[, colSums(!is.na(data)) > 0]

  #define how many columns there are using the ncol() function after deleteing all the empty colmns 
  num_columns <- ncol(data)

  #extract wavelength column 
  wavelength <- data$Wavelength
  
  #create a mapping of first letters to labels
  label_mapping <- setNames(c('0mM', '200mM', '500mM','0mM', '200mM', '500mM'), c('A', 'B', 'C', 'F', 'G', 'H'))
  
  #define a file name to allow for easy naming of the files later on 
  data_file_name_mapping <- setNames(c('1G_','1G_','1G_','4B_','4B_','4B_'), c('A', 'B', 'C', 'F', 'G', 'H'))
  
  #create mapping for pH values based off of second number in column title 
  ph_mapping <- setNames(c('4', '5', '6','7', '8', '9'), c('1', '2', '3', '4', '5', '6'))


  #------------------------------------FOR LOOP----------------------------------------

  #iterate over the columns starting from the second one
  for (i in 2:num_columns) {
    
    #get the column name to determine the label.  
    col_name <- names(data)[i]
    
    #create a new dataframe with the current column as absorbance (wavelength define above)
    current_data <- data.frame(Wavelength = wavelength, data[[i]])
    
    
                 #------------MAPPING CODE-------------------------
    #create variable first_letter to store the first letter of the column name 
    first_letter <- substr(col_name, 1, 1)
    
    #create variable to store the second letter of the column name 
    second_letter <- substr(col_name, 2, 2)
  
    #get the label based on the first letter
    label <- ifelse(first_letter %in% names(label_mapping), label_mapping[[first_letter]], NA)
    
    #get name based on first letter 
    name <- ifelse(first_letter %in% names(data_file_name_mapping), data_file_name_mapping[[first_letter]], NA)
    
    #get pH value based of second string 
    pH <- ifelse(second_letter %in% names(ph_mapping), ph_mapping[[second_letter]], NA)
  
    #construct the file name using the dynamic label
    file_name <- paste0(name, label, "_NaCl_pH",pH,"_Phosphate Buffer", ".xlsx") #MAKE SURE SALT IS CORRECT 
               #------------------------------------------------------    
    
    
    #create dataframe and save as .xlsx (write_xlsx() is a function from the writexl library)
    current_data <- data.frame(Wavelength = wavelength, Absorbance = data[[i]])
    write_xlsx(current_data, path = file_name)
    
    
    #console output 
    message("File: ", file_name, " has been created.")
  
    
  } #END FOR LOOP 

} #END FUNCTION 

```

Run the Script:

```{r}
#THIS SCRIPT WILL SAVE FILES TO YOUR WORKING DIRECTORY 
#MAKE SURE TO RUN THE ABOVE CELL BEFORE TRYING TO RUN THIS CELL 


#MAKE SURE TO CHANE THIS FILE PATH TO YOUR LOCAL DATA FILE PATH 
process_data2('/Users/jackpetersen/Downloads/1G+4B - pH Plate .xlsx')
```

## Original Code: survey1_process_data.R

```{r}
library(tidyverse)
library(googledrive)
library(googlesheets4)
library(writexl)
library(dplyr)

survey1_process_data <- function(input, raw_data, drive_folder_id) {
  
  #function to average data
  #in the future, we may be able to build this into the main function that way we aren't embedding functions, but for right now, this works well
  avg_data <- function(data) {
    
    
    # delete temperature row (can be adjusted later if needed)
    data <- data[, -which(names(data) == "Temperature(¡C)")]
    
    # delete columns with all NA values for all rows (delete empty wells)
    data <- data[, colSums(!is.na(data)) > 0]
    
    # initialize a new dataframe to hold the averaged data
    # keep the first column intact (assuming it's 'Wavelength' or similar)
    averaged_data <- data.frame(Wavelength = data[[1]])
    
    # get all column names except the first one (Wavelength column stays the same)
    column_names <- colnames(data)[-1]
    
    # loop through the columns in sets of three, averaging and adding to the dataset 
    for (i in seq(1, length(column_names), by = 3)) {
      
      # ensure we have a group of three columns
      if (i + 2 <= length(column_names)) {
        
        # group 3 columns together 
        cols_to_average <- column_names[i:(i+2)]
        
        # name the averaged column after all three original columns + _avg
        averaged_column_name <- paste(cols_to_average, collapse = "_") %>% paste0("_avg")
        
        # calculate the mean of the three columns and add it as a new column to 'averaged_data'
        averaged_data[[averaged_column_name]] <- rowMeans(data[, cols_to_average], na.rm = TRUE)
      }
    }
    
    
    
    # print success message to console
    print("Success! Your processed data is now in the dataframe 'averaged_data'")
    
    return(averaged_data)
    
    
  } #end of avg_data 
  
  #-----------------main processing function-----------------------------
  
  
  # average the data. function define above 
  averaged_data <- avg_data(raw_data)
  
  #define salt concentrations and their corresponding column prefixes
  salt_concentrations <- c('0mM' = 'A', '200mM' = 'B', '500mM' = 'C')
  
  #define the column ranges for each peptoid
  peptoid_column_ranges <- list('1' = 1:3, '2' = 5:7, '3' = 9:11, '4' = 1:3, '5' = 5:7, '6' = 9:11)
  
  
 
  for (i in 1:6) {
    peptoid_name <- input[[paste0("peptoid_name_", i)]]
    num_salt_conc <- input[[paste0("num_dif_salt_conc", i)]]
    file_salt_concentrations <- if (num_salt_conc == 1) c("0mM") else names(salt_concentrations)
    
    for (salt_conc in file_salt_concentrations) {
      prefix <- salt_concentrations[[salt_conc]]
      column_range <- peptoid_column_ranges[[as.character(i)]]
      if (i > 3) { prefix <- 'E' }  # For peptoids 4, 5, 6, use 'E' instead of 'A'
      
      # construct the column names for this peptoid and salt concentration
      relevant_column_names <- paste0(prefix, column_range, "_.*_avg")
      
      # find the column name that matches the pattern
      relevant_column <- grep(relevant_column_names, names(averaged_data), value = TRUE)
      
      if (length(relevant_column) == 1) {
        # filter data for the current peptoid and salt concentration
        filtered_data <- averaged_data %>%
          select(Wavelength, Absorbance = all_of(relevant_column))
        
        # construct file name based on user inputs
        file_name <- paste(peptoid_name, salt_conc, input$salt_type, input$pH_tested, input$buffer_type, sep = "_")
        file_name <- paste0(file_name, ".xlsx")
        
        # write the filtered data to an Excel file
        output_file_path <- paste0("/Users/jackpetersen/Desktop/code/AuNP-Peptoid/data", file_name)
        write_xlsx(filtered_data, output_file_path)
        
        # define the Google Drive folder IDs --- MAKE THIS BASED ON NANOPARTICLE SIZE IN FUTURE 
        data_drive_folder_id <- '16VeTz69DEirMMgS6sFvBJVqaW7Lax7cn'

        # upload the file to Google Drive
        drive_upload(media = output_file_path, path = as_id(data_drive_folder_id), name = file_name, overwrite = TRUE)
        
        
        
        #delete the local file after uploading
        if (file.exists(output_file_path)) {
          file.remove(output_file_path)
        }
        
        
        print(file_name)
      } else {
        print(paste0("No data found for ", peptoid_name, " at ", salt_conc))
      }
    }
  }
}

```
